Progress Report

### 2023-09-27 Docs/ProgressReport.Rmd
try splines interpolation of mean and eigenfunctions

library(here)
library(tidyverse)
library(ggpubr)
library(refund)
library(lme4)
library(ROCR)
library(GLMMadaptive)
library(kableExtra)
library(mvtnorm)
theme_set(theme_minimal())

@@

### Debias?

>>>>

library(here)
library(tidyverse)
library(ggpubr)
library(refund)
library(lme4)
library(ROCR)
library(GLMMadaptive)
library(kableExtra)
library(mvtnorm)
library(mgcv)
library(splines)
theme_set(theme_minimal())

@@

### Debias (step 4)

According to the fGFPCA paper, in this step we do two things:

1. Interpolate the eigenfunctions with B-spline basis back to the original grid
2. Debias the eigenvalues with GLMM (mgcv::bam, mgcm::gamm or gamm4:gamm) to be used for Laplace approximation
- we do not need to debias individual scores because the scores for training sample will not be used for out-of-sample prediction

First I'd like to try to fit the debias GLMM model on the full training set with un-interpolated eigenfunctions. The model failed because of vector memory exhaustion. Shall I: 
- Decrease the density of grid (increase bin width)
- Increase memory (re-set to 100GB). But still took a lot of time (more than one hour). The model actually did not finish fitting.

The bias was caused by the misspecification of latent process (assume constant effect over smooth function). And what needs re-evaluation are eigenvalues. Basis functions and mean functions are not biased.

```{r, eval=FALSE}
df_phi <- data.frame(bin=mid, fpca_mod$efunctions[, 1:4])
colnames(df_phi) <- c("bin", paste0("phi", 1:4))

df_train <- df_train %>% left_join(df_phi, by = "bin")
df_train$id <- as.factor(df_train$id)

# usethis::edit_r_environ()
t1 <- Sys.time()
debias_glmm <- bam(Y ~ s(bin, bs="cr", k=10)+
                     s(id, by=phi1, bs="re")+
                     s(id, by=phi2, bs="re")+
                     s(id, by=phi3, bs="re")+
                     s(id, by=phi4, bs="re"), 
                   family = binomial, data=df_train, method = "fREML")
t2 <- Sys.time()
t_debias <- t2=t1

Then I would like to interpolate eigenfunctions using B-spline basis functions:

- Would you interpolate mean and eigenfunctions separately (in my case, five interpolations)? 
- I did not find the reeval_efunctions function

```{r, fig.height=5, fig.width=5}
# mean
interp_mean <- interpSpline(mid, fpca_mod$mu)
full_mean <- predict(interp_mean, x=1:J)
plot(mid, fpca_mod$mu, ylab = "mean", pch = 20)
lines(full_mean$x, full_mean$y, col = "blue")

# PCs
interp_phi1 <- interpSpline(mid, fpca_mod$efunctions[, 1])
full_phi1 <- predict(interp_phi1, x=1:J)
plot(mid, fpca_mod$efunctions[, 1], ylab = "pc1", pch=20)
lines(full_phi1$x, full_phi1$y, col = "blue")

interp_phi2 <- interpSpline(mid, fpca_mod$efunctions[, 2])
full_phi2 <- predict(interp_phi2, x=1:J)
plot(mid, fpca_mod$efunctions[, 2], ylab = "pc2", pch=20)
lines(full_phi2$x, full_phi2$y, col = "blue")

interp_phi3 <- interpSpline(mid, fpca_mod$efunctions[, 3])
full_phi3 <- predict(interp_phi3, x=1:J)
plot(mid, fpca_mod$efunctions[, 3], ylab = "pc3", pch=20)
lines(full_phi3$x, full_phi3$y, col = "blue")

interp_phi4 <- interpSpline(mid, fpca_mod$efunctions[, 4])
full_phi4 <- predict(interp_phi4, x=1:J)
plot(mid, fpca_mod$efunctions[, 4], ylab = "pc4", pch=20)
lines(full_phi4$x, full_phi4$y, col = "blue")
```
